Scattering from compact objects: \\ Regge poles and the complex angular momentum method

### 2020-05-19 rainbowCAM.tex
v2: Typos corrected and minor changes in the text to match the published version

\title{Scattering from compact objects: \\ Regge poles and the Complex Angular Momentum method}

@@

\affiliation{Consortium for Fundamental Physics,  School of Mathematics and Statistics,
University of Sheffield, Hicks Building, Hounsfield Road, Sheffield S3 7RH, United Kingdom \looseness=-1}
%
\author{Tom Stratton}\email{tstratton1@sheffield.ac.uk}
\affiliation{Consortium for Fundamental Physics,  School of Mathematics and Statistics,
University of Sheffield, Hicks Building, Hounsfield Road, Sheffield S3 7RH, United Kingdom \looseness=-1}
%
\author{Sam R. Dolan}\email{s.dolan@sheffield.ac.uk}
\affiliation{Consortium for Fundamental Physics,  School of Mathematics and Statistics,
University of Sheffield, Hicks Building, Hounsfield Road, Sheffield S3 7RH, United Kingdom \looseness=-1}
%
\begin{abstract}
We calculate the Regge poles of the scattering matrix for a gravitating compact body, for scalar fields and for gravitational waves in the axial sector. For a neutron-star-like body, the spectrum exhibits two distinct branches of poles, labelled surface waves and broad resonances; for ultra-compact objects, the spectrum also includes a finite number of narrow resonances. We show, via a WKB analysis, that the discontinuity of the effective potential at the body's surface determines the imaginary component of the broad-resonance poles.

Next, we examine the role of Regge poles in the time-independent scattering of monochromatic planar waves. We apply Complex Angular Momentum techniques to re-sum the partial wave series for the scattering amplitude, expressing it as a residue series evaluated at poles in the first quadrant, accompanied by a background integral. We compute the scattering cross section at several frequencies, and show precise agreement with the partial-wave calculations. Finally, we show that compact bodies naturally give rise to orbiting, glory, and rainbow-scattering interference effects.
\end{abstract}

@@

The time-independent scattering of planar waves in the gravitational field of a compact body has been studied in some detail since the 1960s \cite{Hildreth1964PhDT64, Matzner:1968, Vishveshwara:1970}. A substantial literature has accumulated on \emph{black hole scattering}, focussing on the canonical scenario of a planar wave of circular frequency $\omega$ and spin $s$ \cite{Chrzanowski:1976jb} impinging upon a black hole of mass $M$ in vacuum \cite{Hildreth1964PhDT64, Matzner:1968, Vishveshwara:1970, Mashhoon:1973zz,Fabbri:1975,Sanchez:1977vz,MatznerRyan1978,Handler:1980un,Matzner:1985rjn,Futterman:1988ni,Andersson:1995vi,Glampedakis:2001cx,Dolan:2006vj,Dolan:2007ut,Dolan:2008kf,Crispino:2009xt,Cotaescu:2014jca,Gussmann:2016mkp}. A dimensionless parameter
$
M \omega = \pi r_g / \lambda
$
encapsulates the ratio of the gravitational radius $r_g = 2GM/c^2$ to the wavelength $\lambda$; herein we adopt geometric units such that $G=c=1$. The long-wavelength ($M \omega \ll 1$), short-wavelength ($M \omega \gg 1$) and intermediate regimes have been studied with a combination of perturbative \cite{DeLogi:1977dp,Dolan:2007ut,Guadagnini:2008ha,Sorge:2015yoa}, semi-classical \cite{Matzner:1985rjn, Anninos:1992ih} and numerical methods. The $s = 0$ (scalar) \cite{Matzner:1968,Sanchez:1977vz,Andersson:1995vi,Glampedakis:2001cx,Leite:2019eis}, $s=1/2$ (fermion) \cite{Dolan:2006vj,Cotaescu:2014jca}, $s=1$ (electromagnetic) \cite{Fabbri:1975, Crispino:2009xt, Crispino:2015gua} and $s=2$ (gravitational) cases \cite{MatznerRyan1978,Handler:1980un,Dolan:2008kf} have all been addressed.

Time-independent scattering by a compact body of radius $R$ with a regular centre, such as a neutron star or white dwarf, has received less attention, by comparison. In such a scenario, an electromagnetic wave will not penetrate far inside the compact body; but on the other hand, a gravitational wave will pass through the body without impediment from the matter distribution. A neutron star is also expected to be effectively transparent to neutrinos. Even in such cases where the coupling to matter is negligible, the incident wave will nevertheless be scattered by the spacetime curvature. Thus, the resulting scattering pattern depends not just on $M\omega$, but also on the internal structure of the body, and on its inverse compactness or \emph{tenuity}, $R/M$. Typical values of the tenuity are: $R/M \sim 6$ for neutron stars, $\sim 1.4-9.4 \times 10^3$ for white dwarfs, $4.7 \times 10^5$ for the Sun, and $1.4 \times 10^9$ for Earth. More speculatively, we shall also consider here the case of (hypothetical) \emph{ultra-compact objects} (UCOs), whose tenuity is bounded from below by the Buchdahl bound of $R / M = 9/4$.

Studies of black hole scattering typically involve the calculation of scattering amplitude(s) via partial-wave expansions.  A powerful approach for resumming these partial-wave expansions is provided by so-called Complex Angular Momentum (CAM) techniques. In the CAM approach, the sum over partial waves is replaced by an integral, and the contour of integration is deformed into the complex-angular-momentum plane such that one collects a sum of residues of simple poles: these are the so-called \emph{Regge poles}.

@@

There is a close relationship between Regge poles and QNM frequencies; both are sets of poles of the scattering matrix, with the latter lying in the complex-frequency plane. The excitation of black hole QNMs leads to distinctive signatures in \emph{time-dependent} scattering scenarios. For example, in the immediate aftermath of a black hole merger, the perturbed black hole returns to a quiescent state via a ``ringdown phase'' in which its gravitational-wave signal is well described as a sum of quasinormal modes \cite{Giesler:2019uxc}. The QNM spectrum of compact bodies have been well-studied by several authors  \cite{Detweiler:1985zz, Kokkotas:1986gd, Chandrasekhar449, Kokkotas:1992ka, Leins:1993zz, Andersson:1995ez, Andersson1996}, and are reviewed in Ref.~\cite{Kokkotas:1999bd}.

@@

The remainder of this paper is organised as follows. Sec.~\ref{SecII} reviews the theory of scalar waves on a spherically-symmetric spacetime of a compact object. Here we describe our model spacetime (\ref{SecIIa}), the effective potentials for wave scattering (\ref{subsec:Potentials}) and the physical boundary conditions that define the $S$-matrix (\ref{subsec:bc}). Sec.~\ref{sec:RP} focusses on the Regge pole spectrum. Here we outline the link between Regge poles and QNMs (\ref{subsec:QNMs_RPs}), we describe the numerical method (\ref{subsec:method}), and we present new numerical results for the spectrum (\ref{subsec:results1}) which exhibits three distinct branches of Regge poles (see e.g.~Fig.~\ref{RP_R226_approx_2Mw_3_6_s_1}). In Sec.~\ref{subsec:WKB} we apply the WKB method to obtain an approximate formula for the ``broad resonance'' branch. Sec.~\ref{sec:CAM} concerns the application of CAM techniques to the calculation of the scattering cross section. Here we review the standard partial-wave expansion of the scattering amplitude (\ref{subsec:partial}), we apply CAM techniques to write this as the sum of a residue series and a background integral (\ref{SecIIc}), and we present a selection of numerical results for $d\sigma/d\Omega$ (\ref{subsec:results2}). We conclude with a discussion in Sec.~\ref{sec:conclusions}. In an appendix, we show which quadrants of the complex $\lambda$-plan contain Regge poles of the $S$-matrix.

@@

Effective potentials for the incompressible model are shown in Fig.~\ref{fig:Veff}, for two cases: (i) a neutron-star model with $R = 6M$, and (ii) a ultra-compact object (UCO \cite{Macedo:2018yoi}) with $R = 2.26M$. % (i.e.~close to the Buchdahl bound of $R/M = 9/4$).
In both cases we observe a discontinuity in $V_{\ell}(r)$ across the star's surface, due to the discontinuity in the density $\rho$ (which implies the $C^0$ property of $h(r)$). The jump in the potential takes opposite signs in the scalar-field and gravitational-wave cases, with
\begin{subequations}
\begin{align}
\Delta V_{\ell}^{(s=0)} &= + \frac{3Mf(R)}{R^3} , \\
\Delta V_{\ell}^{\text{ax}} &= - \frac{3Mf(R)}{R^3} ,
\end{align}
\end{subequations}
where
\begin{equation}
\Delta V_{\ell} \equiv \lim_{\epsilon \rightarrow 0} \left\{ V_{\ell}(R + \epsilon) - V_{\ell}(R - \epsilon) \right\} . \label{def:DeltaV}
\end{equation}

@@

\subsection{Boundary conditions and scattering}\label{subsec:bc}
The modes $\phi_{\omega \ell}$ in Eq.~(\ref{eq:sepvariables}) should have a regular behaviour at the centre of the object ($r = 0$), and inspection of the radial equation (\ref{H_Radial_equation}) shows that
\begin{equation}\label{bc_1_in}
\phi_{\omega  \ell}(r) \, \scriptstyle{\underset{r \to 0}{\sim}} \,
\displaystyle{r^{\ell+1}}.
\end{equation}
At the boundary of the compact object, the potential is $C^0$ and thus the mode is $C^2$.
The asymptotic behaviour of the modes far from the body ($r \to +\infty$, or equivalently $r_\ast \to +\infty$) is
\begin{equation}\label{bc_2_in}
\phi_{\omega  \ell}(r) \scriptstyle{\underset{r_\ast \to +\infty}{\sim}}
\displaystyle{ A^{(-)}_\ell (\omega) e^{-i\omega r_\ast} + A^{(+)}_\ell (\omega) e^{+i\omega r_\ast}}.
\end{equation}
With the complex coefficients $A^{(\pm)}_\ell (\omega)$ we then define the \emph{$S$-matrix elements},
\begin{equation}\label{Matrix_S}
  S_{\ell}(\omega) =  e^{i(\ell+1)\pi} \, \frac{A_{\ell}^{(+)}(\omega)}{A_{\ell}^{(-)}(\omega)}.
\end{equation}
We now consider the poles of $S_{\ell}(\omega)$ in the complex plane.

@@

The quasinormal mode spectrum of spherically-symmetric compact objects has been studied in some detail in Refs.~\cite{Kokkotas:1986gd,Kokkotas:1992ka,Andersson:1997eq,Kokkotas:1999bd,Leins:1993zz}.
Newly-formed neutron stars, the remnants of supernovae collapse, are predicted to pulsate with a large initial energy, and fluid pulsations will generate gravitational waves. In 1967, Thorne and Campolattaro \cite{thorne1967non} classified the fluid modes of a relativistic compact body by analogy with the fluid modes of a Newtonian body, with the addition of a damping time due to the emission of GWs. Two decades later, the subject was examined again \cite{Detweiler:1985zz,Kokkotas:1986gd}, and Kokkotas and Schutz \cite{Kokkotas:1992ka} showed the existence of an additional family of modes, dubbed $w$-modes. These modes are characterised by a negligible excitation of fluid motion, and in the axial sector, by no fluid motion at all. They are highly damped and correspond to excitations of the dynamical perturbed space-time. For a review of (gravitational) quasinormal modes in relativistic stars and black holes, see Ref.~\cite{Kokkotas:1999bd}.

@@

It is important to note that, the \textit{continued fraction method} applies to three-term recurrence relations. Leins~\emph{et al.} and BBF use the method introduced by Leaver in
the case of Reissner-Nordst\"om black holes \cite{Leaver:1990zz} to reduce four-term recurrence relations (as in \eqref{Recurrence_4_terms}) to three-term recurrence relations by using a Gaussian elimination step.

We do not use the Gaussian elimination step. Instead we find solutions of \eqref{Recurrence_4_terms} by adapting the \textit{Hill determinant approach} employed by Majumdar and Panchapakesan~\cite{mp} to solve the black hole recurrence formula of Leaver \cite{Leaver:1985ax}. The nontrivial solutions of \eqref{Recurrence_4_terms} exist when the Hill determinant vanishes:
\begin{equation}
\label{Determinant_Hill_4_termes}
D   =  \begin{vmatrix}
   \beta_0         &  \alpha_0    &        0             &       0                    &          0                &     \ldots          &     \ldots        &    \ldots \\
   \gamma_1        &  \beta_1      &     \alpha_1     &       0                    &          0                &     \ldots          &     \ldots        &    \ldots  \\
    \delta_2       &   \gamma_2     &    \beta_2       &   \alpha_2             &         0                &     \ldots           &    \ldots         &    \ldots  \\
    \vdots         &      \ddots     &     \ddots       &    \ddots                &       \ddots          &     \ddots           &   \ldots          &    \ldots \\
     \vdots        &     \vdots      &     \delta_{n-1}      &    \gamma_{n-1}      &     \beta_{n-1}     &      \alpha_{n-1} &    \ddots         &    \ldots  \\
     \vdots        &  \vdots        &     \vdots        &     \delta_n                &    \gamma_{n}    &     \beta_{n}        &     \alpha_{n}  &    \ddots     \\
     \vdots        &  \vdots        &     \vdots        &     \vdots                &    \ddots            &     \ddots            &     \ddots       &    \ddots
                        \end{vmatrix}
       = 0.
\end{equation}
%
Letting $D_n$ be the determinant of the $n \times n$ sub-matrix of $D$ (with diagonal $\{\beta_0,\beta_1,....\beta_n\}$), by \eqref{Recurrence_4_terms}
%
\begin{equation}
\label{derecurrence_4_termes}
D_n=\beta_n D_{n-1} - \gamma_{n}\alpha_{n-1}D_{n-2} + \delta_n \alpha_{n-1} \alpha_{n-2} D_{n-3} ,
\end{equation}
%
with the initial conditions
 \begin{equation}
 \label{Determinant_initial_conds}
 \begin{split}
    D_0 &=\beta_0, \\
    D_1 &=\beta_1\beta_0-\gamma_1\alpha_0 , \\
    D_2 &=\beta_0(\beta_1 \beta_2 - \alpha_1 \gamma_2)- \alpha_0(\alpha_1 \delta_2 -\gamma_1 \beta_2) .
 \end{split}
 \end{equation}
%
Equivalently, % \moh{we can also use this forme}
\begin{align}
\label{recurrence_Hill_RN_4_termes}
D_n    & =\left(\prod_{k=1}^{n+1} k^2\right) P_{n+1} , \nonumber\\
       &  = 1\times 2^2 \ldots (n-2)^2(n-1)^2n^2(n+1)^2 P_{n+1},
\end{align}
where
\begin{eqnarray}
\label{recurrence_Hill_RN_4_termes_bis}
% \nonumber % Remove numbering (before each equation)
P_n = & & \left(\frac{\beta_{n-1}}{n^2}\right)P_{n-1}-\left(\frac{\gamma_{n-1}}{n^2}\right)\left(\frac{\alpha_{n-2}}{(n-1)^2}\right)P_{n-2} \nonumber \\
  &+& \left(\frac{\delta_{n-1}}{n^2}\right) \left(\frac{\alpha_{n-2}}{(n-1)^2}\right) \left(\frac{\alpha_{n-3}}{(n-2)^2}\right) P_{n-3}
,
\end{eqnarray}
with the initial conditions
 \begin{equation}
 \begin{split}
    P_1&=\beta_0 , \\
    P_2&=\frac{\beta_0\beta_1-\gamma_1\alpha_0}{4} ,\\
    P_3&=\frac{\beta_0(\beta_1 \beta_2 - \alpha_1 \gamma_2)- \alpha_0(\alpha_1 \delta_2 -\gamma_1 \beta_2)}{36} .
 \end{split}
 \end{equation}

@@

 \subsection{Numerical results: the spectrum}\label{subsec:results1}
In this section we show numerical results for the Regge pole spectrum of a compact body, in two particular cases: (i) a neutron-star-like body with tenuity $R / M = 6$, and (ii) a UCO with tenuity $R / M = 2.26$, close to the Buchdahl bound. Results for the frequencies $M \omega = 3/2$ and $M \omega = 8$ are compared, for the scalar-field ($s=0$) and axial GW ($s=2$) cases.

Figure \ref{fig:rp1} shows the Regge pole spectrum for a neutron-star-like body ($R/M = 6$). We see that there are two branches of Regge poles in the first quadrant, which meet near $\text{Re}( \lambda_n(\omega) ) \sim \omega b_0$, where $b_0$ is the impact parameter for the null ray which grazes the surface of the compact body. The axial GW $s=2$ modes (filled markers) are typically close to their scalar-field $s=0$ counterparts (unfilled markers), as might be anticipated from the similarities in the effective potentials for the two cases (see Fig.~\ref{fig:Veff}).

@@

Exploring the second and third quadrants, where $\text{Re}[{\lambda}] < 0$, requires a choice to be made on the form of the boundary condition imposed at $r=0$. A natural choice is to impose regularity, i.e.,~$\phi_{\omega\ell} \sim r^{-\lambda+1/2}$ as $r \rightarrow 0$ for $\text{Re} [\lambda] < 0$. An advantage of this choice is that it leads to the same symmetry as in the black hole case for the $S$-matrix elements (see Ref.~\cite{Andersson:1994rk,Folacci:2019cmc}), viz.,
\begin{equation}\label{Matrix_S_CAM_symm}
S_{-\lambda -1/2}(\omega) =  e^{-2i \pi \lambda} \, S_{\lambda -1/2}(\omega)
\end{equation}

@@

The overtones are labelled by $n =1,2,\ldots$ and the condition $\text{Re} \, \lambda_n > 0$ leads to an upper limit for $n$ of %
\begin{equation}\label{Limit_Broad_RP}
n \leq \left\lfloor \frac{\omega R_*}{\pi} \mp \frac{1}{4}\right\rfloor.
\end{equation}
%
In other words, there are a finite number of the broad resonances in the first quadrant.

@@

For any stellar model (such as a polytropic fluid sphere), the potential may in general be $C^{N}$ at the surface (and $C^{\infty}$ elsewhere). In this case, $\mathcal{R}$ and thus the broad resonance branch depends on the discontinuity in the $N$-th derivative of the potential at $r=R$. It can be shown that the imaginary part of the poles is proportional to $(N+1)$ (see Eq.~4.5 of Ref. \cite{Zhang:2011pq}), so in the large-$N$ limit the associated modes have large imaginary parts and little physical consequence.

>>>>

\title{Scattering from compact objects: \\ Regge poles and the complex angular momentum method}

@@

\affiliation{Consortium for Fundamental Physics,  School of Mathematics and Statistics,
University of Sheffield, Hicks Building, Hounsfield Road, Sheffield S3 7RH, United Kingdom \looseness=-1}
%
\author{Tom Stratton}\email{tstratton1@sheffield.ac.uk}
\affiliation{Consortium for Fundamental Physics,  School of Mathematics and Statistics,
University of Sheffield, Hicks Building, Hounsfield Road, Sheffield S3 7RH, United Kingdom \looseness=-1}
%
\author{Sam R. Dolan}\email{s.dolan@sheffield.ac.uk}
\affiliation{Consortium for Fundamental Physics,  School of Mathematics and Statistics,
University of Sheffield, Hicks Building, Hounsfield Road, Sheffield S3 7RH, United Kingdom \looseness=-1}
%
\begin{abstract}
We calculate the Regge poles of the scattering matrix for a gravitating compact body, for scalar fields and for gravitational waves in the axial sector. For a neutron-starlike body, the spectrum exhibits two distinct branches of poles, labeled surface waves and broad resonances; for ultracompact objects, the spectrum also includes a finite number of narrow resonances. We show, via a WKB analysis, that the discontinuity of the effective potential at the body's surface determines the imaginary component of the broad-resonance poles.

Next, we examine the role of Regge poles in the time-independent scattering of monochromatic planar waves. We apply complex angular momentum techniques to re-sum the partial wave series for the scattering amplitude, expressing it as a residue series evaluated at poles in the first quadrant, accompanied by a background integral. We compute the scattering cross section at several frequencies, and show precise agreement with the partial-wave calculations. Finally, we show that compact bodies naturally give rise to orbiting, glory, and rainbow-scattering interference effects.
\end{abstract}

@@

The time-independent scattering of planar waves in the gravitational field of a compact body has been studied in some detail since the 1960s \cite{Hildreth1964PhDT64, Matzner:1968, Vishveshwara:1970}. A substantial literature has accumulated on \emph{black hole scattering}, focusing on the canonical scenario of a planar wave of circular frequency $\omega$ and spin $s$ \cite{Chrzanowski:1976jb} impinging upon a black hole of mass $M$ in vacuum \cite{Hildreth1964PhDT64, Matzner:1968, Vishveshwara:1970, Mashhoon:1973zz,Fabbri:1975,Sanchez:1977vz,MatznerRyan1978,Handler:1980un,Matzner:1985rjn,Futterman:1988ni,Andersson:1995vi,Glampedakis:2001cx,Dolan:2006vj,Dolan:2007ut,Dolan:2008kf,Crispino:2009xt,Cotaescu:2014jca,Gussmann:2016mkp}. A dimensionless parameter
$
M \omega = \pi r_g / \lambda
$
encapsulates the ratio of the gravitational radius $r_g = 2GM/c^2$ to the wavelength $\lambda$; herein we adopt geometric units such that $G=c=1$. The long-wavelength ($M \omega \ll 1$), short-wavelength ($M \omega \gg 1$) and intermediate regimes have been studied with a combination of perturbative \cite{DeLogi:1977dp,Dolan:2007ut,Guadagnini:2008ha,Sorge:2015yoa}, semiclassical \cite{Matzner:1985rjn, Anninos:1992ih} and numerical methods. The $s = 0$ (scalar) \cite{Matzner:1968,Sanchez:1977vz,Andersson:1995vi,Glampedakis:2001cx,Leite:2019eis}, $s=1/2$ (fermion) \cite{Dolan:2006vj,Cotaescu:2014jca}, $s=1$ (electromagnetic) \cite{Fabbri:1975, Crispino:2009xt, Crispino:2015gua} and $s=2$ (gravitational) cases \cite{MatznerRyan1978,Handler:1980un,Dolan:2008kf} have all been addressed.

Time-independent scattering by a compact body of radius $R$ with a regular center, such as a neutron star or white dwarf, has received less attention, by comparison. In such a scenario, an electromagnetic wave will not penetrate far inside the compact body; but on the other hand, a gravitational wave will pass through the body without impediment from the matter distribution. A neutron star is also expected to be effectively transparent to neutrinos. Even in such cases where the coupling to matter is negligible, the incident wave will nevertheless be scattered by the spacetime curvature. Thus, the resulting scattering pattern depends not just on $M\omega$, but also on the internal structure of the body, and on its inverse compactness or \emph{tenuity}, $R/M$. Typical values of the tenuity are: $R/M \sim 6$ for neutron stars, $\sim 1.4-9.4 \times 10^3$ for white dwarfs, $4.7 \times 10^5$ for the Sun, and $1.4 \times 10^9$ for Earth. More speculatively, we shall also consider here the case of (hypothetical) \emph{ultracompact objects} (UCOs), whose tenuity is bounded from below by the Buchdahl bound of $R / M = 9/4$.

Studies of black hole scattering typically involve the calculation of scattering amplitude(s) via partial-wave expansions.  A powerful approach for resumming these partial-wave expansions is provided by so-called complex angular momentum (CAM) techniques. In the CAM approach, the sum over partial waves is replaced by an integral, and the contour of integration is deformed into the complex-angular-momentum plane such that one collects a sum of residues of simple poles: these are the so-called \emph{Regge poles}.

@@

There is a close relationship between Regge poles and QNM frequencies; both are sets of poles of the scattering matrix, with the latter lying in the complex-frequency plane. The excitation of black hole QNMs leads to distinctive signatures in \emph{time-dependent} scattering scenarios. For example, in the immediate aftermath of a black hole merger, the perturbed black hole returns to a quiescent state via a ``ringdown phase'' in which its gravitational-wave signal is well described as a sum of quasinormal modes \cite{Giesler:2019uxc}. The QNM spectrum of compact bodies have been well studied by several authors  \cite{Detweiler:1985zz, Kokkotas:1986gd, Chandrasekhar449, Kokkotas:1992ka, Leins:1993zz, Andersson:1995ez, Andersson1996}, and are reviewed in Ref.~\cite{Kokkotas:1999bd}.

@@

The remainder of this paper is organised as follows. Sec.~\ref{SecII} reviews the theory of scalar waves on a spherically symmetric spacetime of a compact object. Here we describe our model spacetime (\ref{SecIIa}), the effective potentials for wave scattering (\ref{subsec:Potentials}) and the physical boundary conditions that define the $S$-matrix (\ref{subsec:bc}). Sec.~\ref{sec:RP} focusses on the Regge pole spectrum. Here we outline the link between Regge poles and QNMs (\ref{subsec:QNMs_RPs}), we describe the numerical method (\ref{subsec:method}), and we present new numerical results for the spectrum (\ref{subsec:results1}) which exhibits three distinct branches of Regge poles (see e.g.~Fig.~\ref{RP_R226_approx_2Mw_3_6_s_1}). In Sec.~\ref{subsec:WKB} we apply the WKB method to obtain an approximate formula for the ``broad resonance'' branch. Sec.~\ref{sec:CAM} concerns the application of CAM techniques to the calculation of the scattering cross section. Here we review the standard partial-wave expansion of the scattering amplitude (\ref{subsec:partial}), we apply CAM techniques to write this as the sum of a residue series and a background integral (\ref{SecIIc}), and we present a selection of numerical results for $d\sigma/d\Omega$ (\ref{subsec:results2}). We conclude with a discussion in Sec.~\ref{sec:conclusions}. In an appendix, we show which quadrants of the complex $\lambda$-plan contain Regge poles of the $S$-matrix.

@@

Effective potentials for the incompressible model are shown in Fig.~\ref{fig:Veff}, for two cases: (i) a neutron-star model with $R = 6M$, and (ii) a ultracompact object (UCO \cite{Macedo:2018yoi}) with $R = 2.26M$. % (i.e.~close to the Buchdahl bound of $R/M = 9/4$).
In both cases we observe a discontinuity in $V_{\ell}(r)$ across the star's surface, due to the discontinuity in the density $\rho$ (which implies the $C^0$ property of $h(r)$). The jump in the potential takes opposite signs in the scalar-field and gravitational-wave cases, with
\begin{subequations}
\begin{align}
\Delta V_{\ell}^{(s=0)} &= + \frac{3Mf(R)}{R^3} , \\
\Delta V_{\ell}^{\text{ax}} &= - \frac{3Mf(R)}{R^3} ,
\end{align}
\end{subequations}
where
\begin{equation}
\Delta V_{\ell} \equiv \lim_{\epsilon \rightarrow 0} \left\{ V_{\ell}(R + \epsilon) - V_{\ell}(R - \epsilon) \right\} . \label{def:DeltaV}
\end{equation}

@@

\subsection{Boundary conditions and scattering}\label{subsec:bc}
The modes $\phi_{\omega \ell}$ in Eq.~(\ref{eq:sepvariables}) should have a regular behaviour at the center of the object ($r = 0$), and inspection of the radial equation (\ref{H_Radial_equation}) shows that
\begin{equation}\label{bc_1_in}
\phi_{\omega  \ell}(r) \, \scriptstyle{\underset{r \to 0}{\sim}} \,
\displaystyle{r^{\ell+1}}.
\end{equation}
At the boundary of the compact object, the potential is $C^0$ and thus the mode is $C^2$.
The asymptotic behaviour of the modes far from the body ($r \to +\infty$, or equivalently $r_\ast \to +\infty$) is
\begin{equation}\label{bc_2_in}
\phi_{\omega  \ell}(r) \scriptstyle{\underset{r_\ast \to +\infty}{\sim}}
\displaystyle{ A^{(-)}_\ell (\omega) e^{-i\omega r_\ast} + A^{(+)}_\ell (\omega) e^{+i\omega r_\ast}}.
\end{equation}
With the complex coefficients $A^{(\pm)}_\ell (\omega)$ we then define the \emph{$S$-matrix elements},
\begin{equation}\label{Matrix_S}
  S_{\ell}(\omega) =  e^{i(\ell+1)\pi} \, \frac{A_{\ell}^{(+)}(\omega)}{A_{\ell}^{(-)}(\omega)}.
\end{equation}
We now consider the poles of $S_{\ell}(\omega)$ in the complex plane.

@@

The quasinormal mode spectrum of spherically symmetric compact objects has been studied in some detail in Refs.~\cite{Kokkotas:1986gd,Kokkotas:1992ka,Andersson:1997eq,Kokkotas:1999bd,Leins:1993zz}.
Newly formed neutron stars, the remnants of supernovae collapse, are predicted to pulsate with a large initial energy, and fluid pulsations will generate gravitational waves. In 1967, Thorne and Campolattaro \cite{thorne1967non} classified the fluid modes of a relativistic compact body by analogy with the fluid modes of a Newtonian body, with the addition of a damping time due to the emission of GWs. Two decades later, the subject was examined again \cite{Detweiler:1985zz,Kokkotas:1986gd}, and Kokkotas and Schutz \cite{Kokkotas:1992ka} showed the existence of an additional family of modes, dubbed $w$-modes. These modes are characterised by a negligible excitation of fluid motion, and in the axial sector, by no fluid motion at all. They are highly damped and correspond to excitations of the dynamical perturbed space-time. For a review of (gravitational) quasinormal modes in relativistic stars and black holes, see Ref.~\cite{Kokkotas:1999bd}.

@@

It is important to note that, the \textit{continued fraction method} applies to three-term recurrence relations. Leins~\emph{et al.} and BBF use the method introduced by Leaver in
the case of Reissner-Nordst\"om black holes \cite{Leaver:1990zz} to reduce four-term recurrence relations [as in \eqref{Recurrence_4_terms}] to three-term recurrence relations by using a Gaussian elimination step.

We do not use the Gaussian elimination step. Instead we find solutions of \eqref{Recurrence_4_terms} by adapting the \textit{Hill determinant approach} employed by Majumdar and Panchapakesan~\cite{mp} to solve the black hole recurrence formula of Leaver \cite{Leaver:1985ax}. The nontrivial solutions of \eqref{Recurrence_4_terms} exist when the Hill determinant vanishes:
\begin{equation}
\label{Determinant_Hill_4_termes}
D   =  \begin{vmatrix}
   \beta_0         &  \alpha_0    &        0             &       0                    &          0                &     \ldots          &     \ldots        &    \ldots \\
   \gamma_1        &  \beta_1      &     \alpha_1     &       0                    &          0                &     \ldots          &     \ldots        &    \ldots  \\
    \delta_2       &   \gamma_2     &    \beta_2       &   \alpha_2             &         0                &     \ldots           &    \ldots         &    \ldots  \\
    \vdots         &      \ddots     &     \ddots       &    \ddots                &       \ddots          &     \ddots           &   \ldots          &    \ldots \\
     \vdots        &     \vdots      &     \delta_{n-1}      &    \gamma_{n-1}      &     \beta_{n-1}     &      \alpha_{n-1} &    \ddots         &    \ldots  \\
     \vdots        &  \vdots        &     \vdots        &     \delta_n                &    \gamma_{n}    &     \beta_{n}        &     \alpha_{n}  &    \ddots     \\
     \vdots        &  \vdots        &     \vdots        &     \vdots                &    \ddots            &     \ddots            &     \ddots       &    \ddots
                        \end{vmatrix}
       = 0.
\end{equation}
%
Letting $D_n$ be the determinant of the $n \times n$ submatrix of $D$ (with diagonal $\{\beta_0,\beta_1,....\beta_n\}$), by \eqref{Recurrence_4_terms}
%
\begin{equation}
\label{derecurrence_4_termes}
D_n=\beta_n D_{n-1} - \gamma_{n}\alpha_{n-1}D_{n-2} + \delta_n \alpha_{n-1} \alpha_{n-2} D_{n-3} ,
\end{equation}
%
with the initial conditions
 \begin{equation}
 \label{Determinant_initial_conds}
 \begin{split}
    D_0 &=\beta_0, \\
    D_1 &=\beta_1\beta_0-\gamma_1\alpha_0 , \\
    D_2 &=\beta_0(\beta_1 \beta_2 - \alpha_1 \gamma_2)- \alpha_0(\alpha_1 \delta_2 -\gamma_1 \beta_2) .
 \end{split}
 \end{equation}
%
Equivalently, % \moh{we can also use this forme}
\begin{align}
\label{recurrence_Hill_RN_4_termes}
D_n    & =\left(\prod_{k=1}^{n+1} k^2\right) P_{n+1} , \nonumber\\
       &  = 1\times 2^2 \ldots (n-2)^2(n-1)^2n^2(n+1)^2 P_{n+1},
\end{align}
where
\begin{eqnarray}
\label{recurrence_Hill_RN_4_termes_bis}
% \nonumber % Remove numbering (before each equation)
P_n = & & \left(\frac{\beta_{n-1}}{n^2}\right)P_{n-1}-\left(\frac{\gamma_{n-1}}{n^2}\right)\left(\frac{\alpha_{n-2}}{(n-1)^2}\right)P_{n-2} \nonumber \\
  &+& \left(\frac{\delta_{n-1}}{n^2}\right) \left(\frac{\alpha_{n-2}}{(n-1)^2}\right) \left(\frac{\alpha_{n-3}}{(n-2)^2}\right) P_{n-3}
,
\end{eqnarray}
with the initial conditions
 \begin{equation}
 \begin{split}
    P_1&=\beta_0 , \\
    P_2&=\frac{\beta_0\beta_1-\gamma_1\alpha_0}{4} ,\\
    P_3&=\frac{\beta_0(\beta_1 \beta_2 - \alpha_1 \gamma_2)- \alpha_0(\alpha_1 \delta_2 -\gamma_1 \beta_2)}{36} .
 \end{split}
 \end{equation}

@@

 \subsection{Numerical results: The spectrum}\label{subsec:results1}
In this section we show numerical results for the Regge pole spectrum of a compact body, in two particular cases: (i) a neutron-starlike body with tenuity $R / M = 6$, and (ii) a UCO with tenuity $R / M = 2.26$, close to the Buchdahl bound. Results for the frequencies $M \omega = 3/2$ and $M \omega = 8$ are compared, for the scalar-field ($s=0$) and axial GW ($s=2$) cases.

Figure \ref{fig:rp1} shows the Regge pole spectrum for a neutron-starlike body ($R/M = 6$). We see that there are two branches of Regge poles in the first quadrant, which meet near $\text{Re}( \lambda_n(\omega) ) \sim \omega b_0$, where $b_0$ is the impact parameter for the null ray which grazes the surface of the compact body. The axial GW $s=2$ modes (filled markers) are typically close to their scalar-field $s=0$ counterparts (unfilled markers), as might be anticipated from the similarities in the effective potentials for the two cases (see Fig.~\ref{fig:Veff}).

@@

Exploring the second and third quadrants, where $\text{Re}[{\lambda}] < 0$, requires a choice to be made on the form of the boundary condition imposed at $r=0$. A natural choice is to impose regularity, i.e.,~$\phi_{\omega\ell} \sim r^{-\lambda+1/2}$ as $r \rightarrow 0$ for $\text{Re} [\lambda] < 0$. An advantage of this choice is that it leads to the same symmetry as in the black hole case for the $S$-matrix elements (see Refs.~\cite{Andersson:1994rk,Folacci:2019cmc}), viz.,
\begin{equation}\label{Matrix_S_CAM_symm}
S_{-\lambda -1/2}(\omega) =  e^{-2i \pi \lambda} \, S_{\lambda -1/2}(\omega)
\end{equation}

@@

The overtones are labeled by $n =1,2,\ldots$ and the condition $\text{Re} \, \lambda_n > 0$ leads to an upper limit for $n$ of %
\begin{equation}\label{Limit_Broad_RP}
n \leq \left\lfloor \frac{\omega R_*}{\pi} \mp \frac{1}{4}\right\rfloor.
\end{equation}
%
In other words, there are a finite number of the broad resonances in the first quadrant.

@@

For any stellar model (such as a polytropic fluid sphere), the potential may in general be $C^{N}$ at the surface (and $C^{\infty}$ elsewhere). In this case, $\mathcal{R}$ and thus the broad resonance branch depends on the discontinuity in the $N$-th derivative of the potential at $r=R$. It can be shown that the imaginary part of the poles is proportional to $(N+1)$ (see Eq.~(4.5) of Ref. \cite{Zhang:2011pq}), so in the large-$N$ limit the associated modes have large imaginary parts and little physical consequence.
